---
title: "Average number of murder indicator per city in Rio Grande do Sul"
author: "Matheus Souza"
date: "November 2017"
output:
  pdf_document:
    number_sections: yes
  html_document: default
geometry: margin=1.5in, top=0.5in, bottom=0.5in
---

# Installing Section 
 

```{r message=FALSE, warning= FALSE}
install.packages("ggplot2", repos="http://cran.rstudio.com/")
install.packages("dplyr", repos="http://cran.rstudio.com/")
install.packages("magrittr", repos="http://cran.rstudio.com/")
install.packages("tidyr", repos="http://cran.rstudio.com/")
install.packages("reshape", repos="http://cran.rstudio.com/")
install.packages("readr", repos="http://cran.rstudio.com/")
install.packages("curl", repos="http://cran.rstudio.com/")
```


# Loading all required libraries 


```{r message=FALSE, warning= FALSE}

library(ggplot2)
library(dplyr)
library(magrittr)
library(tidyr)
library(readr)
library(reshape)
```


# Loading the data from online resource

A secondary source is used due to an encoding problem in the original data, the only change made was the csv files encoding.  The original data can be founded in http://dados.rs.gov.br/dataset/fee-indicadores-criminais-homicidio-doloso-103597.

This dataset shows the intentional murder count per city in the Rio Grande do Sul state. The original data is separated into different files by year,  but here we aggregate all data to help with operations.  Finally, we end up with a dataset that has six fields and a total number of 3806 observations. 

The fields:
         Município: the city related to a certain intentional murder count;
              Ibge: an official city code, from national statistics bureau; 
latitude/longitude: geographical references;
        Homi_Count: the intentional murder count; 
               Ano: the year that the statistics refer to.

```{r}

df_list <- list()

for (y in 2002:2016){
  URL = paste0("https://raw.githubusercontent.com/mmsouza/lps/master/Final%20Task/fee-",as.character(y),"-mun-indicadores-criminais-homicidio-doloso-103597.csv")
  aux= y-2001
  df_list[[aux]] <- read_delim(URL, delim=",")
  colnames(df_list[[aux]])[5] <- "Homi_Count"
  df_list[[aux]]<-df_list[[aux]] %>% mutate(ano= y)
  
  }

data <- merge_recurse(df_list)

data$Homi_Count <- as.numeric(data$Homi_Count)

```


Now we can find the average of intentional murders per city and see how this value behave in time. To do that a t-test is used to ensure statistic significance. The confidence level used is 95%.


```{r}

data %>% select(Homi_Count,ano) %>% group_by(ano) %>%  do(b = t.test(.$Homi_Count,conf.level=0.95)) %>% mutate( media = unname(b$estimate) ,maxx = c(b$conf.int[2]) ,minn = c(b$conf.int[1])) %>% ggplot(aes(x=ano, y=media)) + geom_errorbar(aes(ymin=minn, ymax=maxx), width=.1) +geom_point() +scale_y_continuous(limits = c(0,30),breaks = seq(0, 30, by = 1)) + scale_x_continuous(limits = c(2001.8,2017),breaks = seq(2002, 2016, by = 1))


```

We can notice a drop in the average in the last two years, but observing the upper and lower limits we can't ensure statistic significance for any assumption about the averages.

Plotting all murder counts could help to understand better the data.


```{r}
 x <- seq(1,nrow(data))
 data %>% ggplot(aes(x= x ,y=Homi_Count)) + geom_point()

```

In the above graph, we can see that are points that could be considered outliers. These outliers don't follow the same pattern of the other points, remove outliers could lead to a more conclusive analysis.


```{r}
data %>% select(Homi_Count,ano) %>% group_by(ano) %>% filter(Homi_Count < 200) %>%  do(b = t.test(.$Homi_Count,conf.level=0.95)) %>% mutate( media = unname(b$estimate) ,maxx = c(b$conf.int[2]) ,minn = c(b$conf.int[1])) %>% ggplot(aes(x=ano, y=media)) + geom_errorbar(aes(ymin=minn, ymax=maxx), width=.1) +geom_point() +scale_y_continuous(limits = c(0,13),breaks = seq(0, 13, by = 1)) + scale_x_continuous(limits = c(2001.8,2017),breaks = seq(2002, 2016, by = 1))

```

Now we found a significant reduction in the average of intentional murders in the years 2015 and 2016.

However, until now data integrity and coherence wasn't explored. First, let's see the number of observations in each year.  


```{r}

data %>% group_by(ano) %>% summarise(tam=n()) %>% arrange(desc(ano))

```

Now we can see the reason for that significant drop in the average, the last two years observations count are more than twice the others years. Beyond the unbalanced observation count, is important to know if cities observed are coherent.  For instance, in 2014 and 2008 both have an observation count of 223, but are all these 223 cities the same?

```{r}
df2014 <- data %>% filter(ano == 2014)

data %>% filter(ano == 2008) %>%  filter(Município %in% df2014$Município )  %>% summarise(tam=n())
```

we found that only 155 of 223 observations are about the same city.  In order to maintain the coherence, we should find the sub-dataset that has exactly the same cities. 

```{r}


data %>% group_by(Município) %>% filter(n() == 15)  %>%  group_by(ano) %>%  summarise(tam=n())
sub_data <- data %>% group_by(Município) %>% filter(n() == 15)
```

Only 63 cities are present in all 15 years of data collections, we can calculate the averages again now over this reduced sub-dataset.

```{r}
sub_data %>% select(Homi_Count,ano) %>% group_by(ano) %>%  do(b = t.test(.$Homi_Count,conf.level=0.95)) %>% mutate( media = unname(b$estimate) ,maxx = c(b$conf.int[2]) ,minn = c(b$conf.int[1])) %>% ggplot(aes(x=ano, y=media)) + geom_errorbar(aes(ymin=minn, ymax=maxx), width=.1) +geom_point() +scale_y_continuous(limits = c(0,60),breaks = seq(0, 60, by = 4)) + scale_x_continuous(limits = c(2001.8,2017),breaks = seq(2002, 2016, by = 1))

sub_data %>% select(Homi_Count,ano) %>% group_by(ano) %>% filter(Homi_Count < 200) %>% do(b = t.test(.$Homi_Count,conf.level=0.95)) %>% mutate( media = unname(b$estimate) ,maxx = c(b$conf.int[2]) ,minn = c(b$conf.int[1])) %>% ggplot(aes(x=ano, y=media)) + geom_errorbar(aes(ymin=minn, ymax=maxx), width=.1) +geom_point() +scale_y_continuous(limits = c(0,60),breaks = seq(0, 60, by = 4)) + scale_x_continuous(limits = c(2001.8,2017),breaks = seq(2002, 2016, by = 1))

```

In this second analysis both, with and without outliers, point to an inconclusive result. The first, even result in differences over 40 murders between the upper and lower limits.   These results reveal a necessity for a more rigorous conduct on the statistics collecting process and the importance of pay attention to the validation process.


